Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-22 paper.tex
save

\begin{table}
\caption{Average precision, recall and f-measure scores on test folds.}
\label{table:results}
\centering
\begin{tabular}{|l|c c c|c c c|}
  \hline
                       & \multicolumn{3}{|c|}{\textbf{B3}} & \multicolumn{3}{|c|}{\textbf{Pairwise}} \\
  \textbf{Description} & $P$ & $R$ & $F$ & $P$ & $R$ & $F$\\
  \hline
  \hline
Baseline & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
\hline
Blocking = SFI & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
Blocking = Double metaphone & 0.9856 & 0.9827 & 0.9841 & 0.9927 & 0.9817 & 0.9871 \\
Blocking = NYSIIS & 0.9875 & 0.9826 & 0.9850 & 0.9936 & 0.9814 & 0.9875 \\
Blocking = Soundex & 0.9886 & 0.9745 & 0.9815 & 0.9935 & 0.9725 & 0.9828 \\
\hline
No name normalization & 0.9887 & 0.9697 & 0.9791 & 0.9931 & 0.9658 & 0.9793 \\
\hline
Classifier = GBRT & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
Classifier = Random Forests & 0.9909 & 0.9783 & 0.9846 & 0.9957 & 0.9752 & 0.9854 \\
Classifier = Linear Regression & 0.9749 & 0.9584 & 0.9666 & 0.9717 & 0.9569 & 0.9643 \\
\hline
Training pairs = Non-blocked, uniform & 0.9793 & 0.9630 & 0.9711 & 0.9756 & 0.9629 & 0.9692 \\
Training pairs = Blocked, uniform & 0.9854 & 0.9720 & 0.9786 & 0.9850 & 0.9707 & 0.9778 \\
Training pairs = Blocked, balanced & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
\hline
Clustering = Average linkage & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
Clustering = Single linkage & 0.9741 & 0.9603 & 0.9671 & 0.9543 & 0.9626 & 0.9584 \\
Clustering = Complete linkage & 0.9862 & 0.9709 & 0.9785 & 0.9920 & 0.9688 & 0.9803 \\
\hline
No cut & 0.9024 & 0.9828 & 0.9409 & 0.8298 & 0.9776 & 0.8977 \\
Global cut & -- & -- & -- & -- & -- & -- \\
Block cut & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
  \hline
\end{tabular}
\end{table}

\textit{Baseline.} Results presented in this section are all discussed with
respect to a baseline solution using the following combination of components:
\begin{itemize}
\item Blocking: same surname and the same first given name initial strategy (SFI);
\item Linkage function: all 22 features defined in Table~\ref{table:features},
    gradient boosted regression trees as supervised learning algorithm
    and a training set of pairs built by balancing easy and difficult cases.
\item Clustering: agglomerative clustering using average linkage and
    block cuts found to maximize $F_\text{B3}({\cal C}_\text{train}^\prime, \widehat{\cal C}_\text{train}^\prime, {\cal S}_\text{train}^\prime)$.
\end{itemize}

\textit{Blocking.}

\glnote{Discuss baseline + different blocking strategies}\\
\glnote{Discuss blocking in terms of maximum recall}

\begin{table}
\caption{Maximum recall of blocking strategies.}
\label{table:results}
\centering
\begin{tabular}{|l|c|c|}
  \hline
  \textbf{Blocking} & $R_\text{B3}^*$ & $R_\text{pairwise}^*$ \\
  \hline
  \hline
    SFI & 0.9828 & 0.9776 \\
    Double metaphone & 0.9907 & 0.9863 \\
    NYSIIS & 0.9902 & 0.9861 \\
    Soundex & 0.9906 & 0.9863 \\
  \hline
\end{tabular}
\end{table}

% \glnote{To review}

% Let us introduce a term \textit{blocking limited maximum recall} to describe the highest
% recall that the algorithm can theoretically reach after blocking is done.
% Formally, for the set of signatures ${\cal S}$, blocking function $F: \mathcal{S}
% \rightarrow \mathcal{B}$, where $ \mathcal{B} = \{s_{0}, \ldots{} , s_{N-1}\}$ is a
% partition of $\mathcal{S} $ , the recall function $\mathcal{R}: \mathcal{C}_{gold},
% \mathcal{C}_{predicted} \rightarrow \langle0,1\rangle$ where $\mathcal{C}_{gold}$ is the
% partition of $\mathcal{S}$ where for every author there is a set in $\mathcal{C}_{gold}$
% that contains only all the signatures of this author, $\mathcal{C}_{predicted}$ is the
% partition of $\mathcal{S}$ output by the algorithm, and $\mathcal{Q} = \{\cup_{i=0}^{N-1}
% q_i $ : $\forall_{i \in \langle0,N-1\rangle} q_i$ is a partition of $s_i$\}. Then
% \textit{blocking limited maximum recall} is defined as $\max_{q \in \mathcal{Q}}
% \mathcal{R}(\mathcal{C}_{gold}, q)$. For sanely defined recall functions, such as $B^{3}$
% recall or \textit{pairwise} recall \msnote{Further explanation needed, why it is
% equivalent?}, it is equivalent to $\mathcal{R}(\mathcal{C}_{gold}, \mathcal{B})$.

% The \textit{blocking limited maximum recall} for the last name and first initial blocking
% and $B^3$ scoring equals 0.9827. Comparing it to the $B^3$ recall of the base experiment,
% one can see that a huge part of errors comes from the blocking step which forbids some of
% the pairs to be merged together. To prevent that, we tried the phonetic blocking algorithm
% using three different phonetic algorithms: Double Metaphone \citep{doublemetaphone}, NYSIIS
% \citep{nysiis} and Soundex \citep{Soundex}. As Double Metaphone can return two results for a
% string, we always discarded the second one when we used this algorithm.

% \msnote{Mention two new features for this experiments, other than that everything is the
% same as in the base experiment}

% compare with torvik & smalheiser (2009) who argue that lnfi blocking has recall around 98%
% MS - recall for lnfi (blocking only) 0.9776, lnfi (whole algorithm) 0.9739,
% MS - all three results use pairwise scoring

>>>>

\begin{table}
\caption{Average precision, recall and f-measure scores on test folds.}
\label{table:results}
\centering
\begin{tabular}{|l|c c c|c c c|}
  \hline
                       & \multicolumn{3}{|c|}{\textbf{B3}} & \multicolumn{3}{|c|}{\textbf{Pairwise}} \\
  \textbf{Description} & $P$ & $R$ & $F$ & $P$ & $R$ & $F$\\
  \hline
  \hline
Baseline & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
\hline
Blocking = SFI & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
Blocking = Double metaphone & 0.9856 & 0.9827 & 0.9841 & 0.9927 & 0.9817 & 0.9871 \\
Blocking = NYSIIS & 0.9875 & 0.9826 & 0.9850 & 0.9936 & 0.9814 & 0.9875 \\
Blocking = Soundex & 0.9886 & 0.9745 & 0.9815 & 0.9935 & 0.9725 & 0.9828 \\
\hline
No name normalization & 0.9887 & 0.9697 & 0.9791 & 0.9931 & 0.9658 & 0.9793 \\
\hline
Classifier = GBRT & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
Classifier = Random Forests & 0.9909 & 0.9783 & 0.9846 & 0.9957 & 0.9752 & 0.9854 \\
Classifier = Linear Regression & 0.9749 & 0.9584 & 0.9666 & 0.9717 & 0.9569 & 0.9643 \\
\hline
Training pairs = Non-blocked, uniform & 0.9793 & 0.9630 & 0.9711 & 0.9756 & 0.9629 & 0.9692 \\
Training pairs = Blocked, uniform & 0.9854 & 0.9720 & 0.9786 & 0.9850 & 0.9707 & 0.9778 \\
Training pairs = Blocked, balanced & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
\hline
Clustering = Average linkage & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
Clustering = Single linkage & 0.9741 & 0.9603 & 0.9671 & 0.9543 & 0.9626 & 0.9584 \\
Clustering = Complete linkage & 0.9862 & 0.9709 & 0.9785 & 0.9920 & 0.9688 & 0.9803 \\
\hline
No cut & 0.9024 & 0.9828 & 0.9409 & 0.8298 & 0.9776 & 0.8977 \\
Global cut \glnote{TODO} & -- & -- & -- & -- & -- & -- \\
Block cut & 0.9901 & 0.9760 & 0.9830 & 0.9948 & 0.9738 & 0.9842 \\
  \hline
\end{tabular}
\end{table}

\textit{Baseline.} Results presented in Table~\ref{table:results} are all discussed with
respect to a baseline solution using the following combination of components:
\begin{itemize}
\item Blocking: same surname and the same first given name initial strategy (SFI);
\item Linkage function: all 22 features defined in Table~\ref{table:features},
    gradient boosted regression trees as supervised learning algorithm
    and a training set of pairs built by balancing easy and difficult cases.
\item Clustering: agglomerative clustering using average linkage and
    block cuts found to maximize $F_\text{B3}({\cal C}_\text{train}^\prime, \widehat{\cal C}_\text{train}^\prime, {\cal S}_\text{train}^\prime)$.
\end{itemize}

\textit{Blocking.} The very good precision of the baseline ($0.9901$) but its
lower recall ($0.9760$) suggests that the blocking strategy might be the
limiting factor to further overall improvements. As Table~\ref{table:blocking}
indeed indicates, the maximum recall (i.e., if within a block, all signatures
were clustered optimally) for SFI is $0.9828$ , hence not letting much more
room for improvement. At the price of fewer and therefore slightly larger
blocks (as reported in the right column of Table~\ref{table:blocking}), the
proposed phonetic-based blocking strategies show better maximum recall (all
around $0.9905$), thereby pushing further the upper bound on the maximum overall
performance of author disambiguation. As Table~\ref{table:results} indeed shows,
switching to either Double metaphone or NYSIIS phonetic-based blocking allows
to improve the overall F-measure score, trading precision for recall. In particular,
the NYSIIS-based phonetic blocking shows to be the most effective when applied
to the baseline (with an F-measure of $0.9850$) while also being the most
efficient computationally (with 10857 blocks versus 12978 for the baseline).

Finally, let us also note that Table~\ref{table:blocking} corroborates the
estimation of \citep{torvik2009author}, stating that SFI blocking has a recall
around $98\%$ on real data.

\begin{table}
\caption{Maximum recall of blocking strategies, and their number of blocks on ${\cal S}^\prime$.}
\label{table:blocking}
\centering
\begin{tabular}{|l|cc|c|}
  \hline
  \textbf{Blocking} & $R_\text{B3}^*$ & $R_\text{pairwise}^*$ & \# blocks \\
  \hline
  \hline
    SFI & 0.9828 & 0.9776 & 12978 \\
    Double metaphone & 0.9907 & 0.9863 & 9753 \\
    NYSIIS & 0.9902 & 0.9861 & 10857 \\
    Soundex & 0.9906 & 0.9863 & 9403 \\
  \hline
\end{tabular}
\end{table}

\textit{Normalization.} As discussed before, the seemingly anodyne step of
normalizing author names (stripping accents, removing affixes), as done in the
baseline, reveals to be quite significant. Results of Table~\ref{table:results}
indeed clearly suggests that not normalizing significantly reduces performance
(yielding an F-measure of $0.9830$ when normalizing, but decreasing to $0.9791$
when raw author name strings are used instead).
